---
title: "lczexplore_alter"
author: "Matthieu Gousseff"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
resource_files:
  - vignettes/Redon.png
vignette: >
 %\VignetteIndexEntry{lczexplore_alter}
 %\VignetteEngine{knitr::knitr}
 %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
        collapse = TRUE,
        comment = "#>"
)
if (!require("png")) {
  install.packages("png")
  library("png")
}

```
The lczexplore package was developed thanks to the Pændora 2 project, funded by ADEME, a French agency for ecological transition.

The package lczexplore allows to group several Local Climate Zone levels (LCZ) into broader categories. The `LCZgroup2` function allows this grouping, and the main functions of the package allow to specify the expected levels and colors to use.  
the lczexplore also allows the comparison of two sets of qualitative variables on the same territory, with the `importQualSurf` function.

# Group levels and explore the resulting broader categories
## From original data... 
Load the package. 

```{r}
library(lczexplore)
```

Load the example data of Redon City.
```{R}
dirPath<-paste0(system.file("extdata",package="lczexplore"),"/")
dirPathOSM<-paste0(dirPath,"osm/2022/Redon")
dirPathBDT<-paste0(dirPath,"bdtopo_2_2/Redon")
redonOSM<-importLCZgen(dirPath=dirPathOSM,file="rsu_lcz.geojson",column = "LCZ_PRIMARY")
redonBDT<-importLCZgen(dirPath=dirPathBDT,file="rsu_lcz.geojson",column = "LCZ_PRIMARY")
#
```
Compare the LCZ classifications produced with GeoClimate and OpenStreetMAp and French BDTOPO data. 

```{R, fig.dim = c(8, 6)}

comparison<-compareLCZ(sf1=redonBDT,column1="LCZ_PRIMARY", wf1="BD TOPO v2.2",
                        sf2=redonOSM,column2="LCZ_PRIMARY",wf2="Open Street Map", ref=1,
                        repr="standard",exwrite=F,location="Redon",saveG="")
```
To see if the urban envelopes differ, one may want to group all the urban LCZ. 


## ... to grouped levels

For instance, one may want to illustrate the contour of a city, by grouping the urban LCZ and the vegetation ones. Use the `LCZgroup2` function and set the dataset with the argument `sf` and the column with `column`.Then you pass as many vectors as there are grouped category, each of them containing the levels to be grouped together. The newly created categories will be stored in the `outCol` column. If the user doesn't enter them, by default the column is called 'grouped'.

The newly created dataset can be mapped with `showLCZ` function.

```{r, fig.dim = c(8, 6)}
redonOSMgrouped<-LCZgroup2(redonOSM,column="LCZ_PRIMARY",
                           urban=c("1","2","3","4","5","6","7","8","9"),
                           industry="10",
                           vegetation=c("101","102","103","104"),
                           impervious="105",pervious="106",water="107",cols=c("red","black","green","grey","burlywood","blue"))
showLCZ(redonOSMgrouped, column="grouped",repr="alter",wf="OSM",
        LCZlevels = c("urban","industry","vegetation","impervious","pervious","water"),
        cols=c("red","black","green","grey","burlywood","blue"))

```

# Compare two LCZ classification with the function compareLCZ

The package intends to make exploring pairs of LCZ maps easy. 
In the `alter` representation, one can specify levels of the data and colors to be associated with these levels in several ways. 
The package will also try to cope with errors and display messages to help users to see what actions were taken. 

## Levels specified but no colors

